Manuscript

### 2019-01-24 manuscript.tex
updated manuscript_analysis and manuscript

We derived our clinical associations from a new project conducted by the
Variant Interpretation for Cancer Consortium (VICC;
\url{https://cancervariants.org/}), a Driver Project for the Global
Alliance for Genomics and Health (GA4GH) (The Global Alliance for
Genomics and Health, 2016). The VICC has curated annotations of known
cancer variants at varying levels of evidence from multiple resources
(A. H. Wagner et al., 2018): the Precision Medicine Knowledgebase (L.
Huang et al., 2016), MolecularMatch
(\url{https://www.molecularmatch.com/index.html}), OncoKB (Chakravarty
et al., 2017), Jackson Labs Clinical Knowledgebase (S. E. Patterson et
al., 2016), Clinical Interpretations of Variance in Cancers (CIViC)
(Griffith et al., 2017), and the Cancer Genome Interpreter (CGI)
(Tamborero et al., 2018). These harmonized variants are hosted on an
ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P,
\url{https://search.cancervariants.org/\#*}) that allows users to query
and filter aggregated evidence from the various databases listed above
as well as the GA4GH beacon service
(\url{http://beacon-network.org/\#/}), which yields access to genetic
mutation data from over 200 datasets as of 2016 (The Global Alliance for
Genomics and Health, 2016). For drug screen datasets, we focus on three
well-known, large-scale drug projects: the CCLE and GDSC, which were
mentioned above, and the Center for the Science of Therapeutics at the
Broad Institute's Cancer Therapeutics Response Portal (CTRP) dataset
(Basu et al., 2013; M. G. Rees et al., 2016; B. Seashore-Ludlow et al.,
2015). We also analyzed the results from a large-scale CRISPR/Cas9
screen conducted by the Broad Institute's Cancer Dependency Map project
(DepMap) using the Avana knockout library (Doench et al., 2016; Meyers
et al., 2017), which will address whether gene essentiality screens via
CRISPR yield results comparable to functional drug screens.

@@

After filtering, 894 level-A G2P interpretations were carried forward in
subsequent analyses, constituting 57 unique drugs and 34 unique genes in
88 distinct level-A combinations (\emph{Figure 1A}). The number of
level-A interpretations per drug ranged from 1 to 176 (\(M =\) 15.68,
\(SD =\) 34.63), 1 to 264 (\(M =\) 26.24, \(SD =\) 61.58) per gene, and
for each level-A drug-gene combination, the range was 1 to 103 (\(M =\)
10.14, \(SD =\) 19.80). These unique drugs, genes, and combinations are
hereafter referred to as G2P drugs, G2P genes, and G2P associations.
Filtering our screening datasets by these clinical criteria yields
panels of potentially clinically relevant omic features. Of the 57
unique G2P drugs, 37 of them were screened in at least one of the drug
screen datasets, 17 drugs were screened in at least two of the three
datasets, and only 3 were screened in all three datasets. The remaining
20 un-screened G2P drugs suggests an incompleteness of these drug
screens, though these projects frequently release new screening data
that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P
associations tested in the CCLE, CTRP, and GDSC drug screens. Of the 88
G2P associations, data was available for 9 associations in CCLE, 48 in
CTRP, and 44 in GDSC. Controlling for a false discovery rate (FDR)
\textless{} 5\%, CCLE yielded no significant correlations (Wilcoxon
rank-sum tests, \(p_k = 2.1^{-4}\)), though the most significant G2P
association was selumetinib-\emph{KRAS} (\(p_{adj} = 1.9^{-3}\)). In
CTRP, AUC from 3 drug-gene associations correlated with the gene's
mutation status (\(p_k = 4.5^{-3}\)): selumetinib-\emph{KRAS}
(\(p_{adj} = 4.4^{-8}\)), vemurafenib-\emph{BRAF}
(\(p_{adj} = 4.4^{-8}\)), and dabrafenib-\emph{BRAF}
(\(p_{adj} = 7.7^{-5}\)). Dabrafenib-\emph{BRAF}
(\(p_{adj} = 8.6^{-11}\)) and trametinib-\emph{BRAF}
(\(p_{adj} = 4.2^{-6}\)) were significant in GDSC (\(p_k = 2.3^{-3}\)).

@@

To test for potentially novel drug-gene associations, we conducted
similar Wilcoxon rank-sum tests for all pairwise combinations of G2P
genes and drugs tested in the three drug screens (\emph{Figures 1B and
2A}). At an unadjusted \(\alpha = 0.05\), the tests revealed that in
CCLE, AUC significantly correlated with mutation status for 18 out of
170 (10.59\%) tested drug-gene combinations were significantly
different, 135 out of 918 (14.71\%) were significantly different for
CTRP, and 89 out of 850 (10.47\%) for GDSC.

Controlling for a FDR \textless{} 5\%, selumetinib-\emph{BRAF}
(\(p_{adj} = 1.4^{-10}\)) was the only significant drug-gene combination
tested in CCLE (\(p_k = 7.9^{-4}\)). CTRP (\(p_k = 6.5^{-4}\)) yielded 5
significant combinations, one of which was selumetinib-\emph{BRAF}.
There were 3 significant G2P associations in the GDSC dataset
(\(p_k = 9.5^{-5}\)). Of all the significant correlations between
mutation status and AUC across datasets, the only genes represented were
\emph{KRAS} and \emph{BRAF} (Figure~\ref{fig:drug-venn}). Of the 7
significant drug-gene associations across datasets, 4 were G2P
associations. The remaining 3 associations were selumetinib-\emph{BRAF},
trametinib-\emph{KRAS}, and pazopanib-\emph{BRAF}.

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/drugscreen_venns.pdf}
\caption{\label{fig:drug-venn}Overlap of drug-gene associations drawn from all
pairwise combinations of G2P drugs and G2P genes whose drug effects
significantly correlated with (\textbf{A}) mutation status, (\textbf{B})
gene expression, and (\textbf{C}) copy number across CCLE, CTRP, and
GDSC.}
\end{figure}

@@

Prior to correction for an FDR of 5\%, no DGLs were identified in the
CCLE screen when using mutation status as a correlate (\(n =\) 7, 2 DGLs
were not analyzed due to lack of mutation annotations in \emph{UGT1A1}
in \enquote{cancer} lineages), but CTRP (\(n =\) 48) and GDSC (\(n =\)
39) identified 6 and 5, respectively, in skin cancer, lung cancer, and
cancer. After Benjamini-Hochberg FDR correction, none of the datasets
yielded significant DGL interactions.

\begin{quote}
\textbf{(???) KS test results}
\end{quote}

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/g2p_spearman_ks_res_plot.pdf}
\caption{\label{fig:ksplot}Comparison of Spearman correlation coefficients (\(r_s\))
derived from copy number (log\(_2\) ratio), gene expression (RPKM), and
AUC z-scores from G2P drug-gene associations. Coefficients were grouped
by whether or not the associated lineage for each comparison was in a
G2P drug-gene-lineage association or not. Distributions of these grouped
coefficients were compared using Kolmogorov-Smirnov tests.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/crispr_venn.pdf}
\caption{\label{fig:crispr-venn}Overlap of genes whose gene expression and/or copy
number significantly correlated with CERES score.}
\end{figure}

@@

When we don't account for lineage, Wilcoxon rank-sum tests
(\(p_k = 5.1^{-3}\), FDR \textless{} 5\%) revealed that mutations in 2
genes significantly correlated with CERES score: \emph{KRAS}
(\(p_{adj} = 2.6^{-37}\)) and \emph{BRAF} (\(p_{adj} = 8.2^{-18}\)).
Gene expression (RPKM) of 12 out of 33 (36.36\%) G2P genes significantly
correlated with CERES score, as did copy number (log2-ratio) for 12 out
of 33 (36.36\%) G2P genes (Figure~\ref{fig:crispr-venn}).

For the purpose of comparison to the drug screen data, if we treat
CRISPR knockdown like a small-molecule drug, then these results are more
comprehensive than the results from CCLE, CTRP, and GDSC (Table 2).

\begin{itemize}
\tightlist
\item
  Lineage?
\end{itemize}

@@

In a lineage-agnostic context of the CRISPR screen, ERBB2 gene
essentiality score had a significant negative correlation with ERBB2
gene expression (\(p < 10^{-12}\), \(r = -0.31\)) and ERBB2 copy number
(\(p < 10^{-15}\), \(r = -0.34\)) (\emph{Figure \#\#2}). Similarly, MET
gene essentiality also correlated with both MET gene expression
(\(p < 10^{-12}\), \(r = -0.31\)) and copy number (\(p < 10^{-8}\),
\(r = -0.25\)). Notably, BRCA1 also significantly correlated with both
gene expression (\(p < 0.001\), \(r = 0.16\)) and copy number
(\(p < 0.01\), \(r = 0.14\)). BRCA2 significantly correlated with all
three of the genomic features tested: gene expression (\(p < 0.05\),
\(r = 0.10\)), copy number (\(p < 10^{-7}\), \(r = 0.23\)), and mutation
status (\(p < 0.001\)).

@@

\textbf{Should we do this?}

@@

CRISPR/Cas9, CCLE mutation calls, and CCLE copy number data were all
retrieved from the Broad Institute's Cancer Dependency Map (DepMap)
Public 18Q4 release via the DepMap data portal
(\url{https://depmap.org/portal/download/}) (Broad DepMap, 2018). Gene
expression data was retrieved from the DepMap Public 18Q3 release (Broad
DepMap, 2018) from the same data portal. For cell lines with no CCLE
annotation, the DepMap group drew raw copy number and mutation data from
whole exome sequencing data produced by the Wellcome Trust Sanger
Institute {[}Catalogue Of Somatic Mutations In Cancer (COSMIC,
\url{https://cancer.sanger.ac.uk/cell_lines}) (S. Bamford et al., 2004;
S. A. Forbes et al., 2017); European Genome-phenome Archive (Lappalainen
et al., 2015), accession number EGAD00001001039{]} and processed the
data following the CCLE pipelines to ensure consistency.

@@

Gene effect scores for 17,634 genes in 517 cell lines were inferred from
a CRISPR/Cas9 (clustered regularly interspaced short palindromic
repeats/CRISPR-associated 9) screen using the Broad Institute's Avana
knockout library (Doench et al., 2016). The Broad Institute's data
processing and screening methods are available from the figshare record
(Broad DepMap, 2018) and the original publication of CERES, the
algorithm that computes inferred gene dependency scores (Meyers et al.,
2017). The DepMap releases new data quarterly, with the current data set
being the 18Q4 release. This release has 175 more cell lines than the
original data release that was published with CERES (dataset:
gene\_effect.csv, accessed 15 November 2018).

@@

The CCLE generated genomic copy number (CN) data using the Affymetrix
Genome-Wide Human SNP Array 6.0 and GenePattern pipeline (The Cancer
Genome Atlas Research Network, 2008) and normalized segmented CN
log2-ratios for 23,299 genes across 1,098 cell lines using circular
binary segmentation (Olshen, Venkatraman, Lucito, \& Wigler, 2004)
(dataset: public\_18Q4\_gene\_cn.csv, accessed 15 November 2018).

@@

For \emph{Figure 1}, all data were imported and manipulated using pandas
(v0.23.0) (McKinney, 2011) and plotted using seaborn (v0.90) (Waskom,
2012), which sat on top of matplotlib (v2.2.2) (Hunter, 2007) and was
executed on Python (v2.7.15) (Python Software Foundation,
\url{https://www.python.org/}) in an IPython (v5.5.0) (Pérez \& Granger,
2007) kernel within a localhost Jupyter (v5.2.2) notebook (Kluyver et
al., 2016). All other analyses were conducted in R (v3.5.0) (R
Development Core Team, 2010, \url{https://www.R-project.org/}) in an
RStudio environment (v1.1.447) (RStudio, 2012). All other plots were
drawn using the ggplot2 (v3.0.0) R package (Wickham, 2016).

@@

For computing Spearman correlation coefficients, we used the
\texttt{cor.test} base R package. If there were 3 or fewer data points,
then the correlation was not done and a value of \texttt{NA} was
supplied.

@@

@@

\hypertarget{ref-broaddepmap2018a}{}
Broad DepMap. (2018, January 11). DepMap Achilles 18Q4 public. Figshare.
Retrieved from
\url{https://figshare.com/articles/DepMap_Achilles_18Q4_public/7270880}

@@

@@

@@

@@

\hypertarget{ref-rdevelopmentcoreteam2010}{}
R Development Core Team. (2010). \emph{R: A Language and Environment for
Statistical Computing}. Vienna, Austria: R Foundation for Statistical
Computing. Retrieved from \url{https://www.R-project.org/}

@@

\hypertarget{ref-thecancergenomeatlasresearchnetwork2008}{}
The Cancer Genome Atlas Research Network. (2008). Comprehensive genomic
characterization defines human glioblastoma genes and core pathways,
\emph{455}, 1061--1068.
doi:\href{https://doi.org/10.1038/nature07385}{10.1038/nature07385}

>>>>

To curate an evidence-based list of these clinical biomarkers, we
utilized a new oncological data aggregation project conducted by the
Variant Interpretation for Cancer Consortium (VICC;
\url{https://cancervariants.org/}), a Driver Project for the Global
Alliance for Genomics and Health (GA4GH) (The Global Alliance for
Genomics and Health, 2016). The VICC has curated annotations of known
cancer variants at varying levels of evidence from multiple resources
(A. H. Wagner et al., 2018): the Precision Medicine Knowledgebase (L.
Huang et al., 2016), MolecularMatch
(\url{https://www.molecularmatch.com/index.html}), OncoKB (Chakravarty
et al., 2017), Jackson Labs Clinical Knowledgebase (S. E. Patterson et
al., 2016), Clinical Interpretations of Variance in Cancers (CIViC)
(Griffith et al., 2017), and the Cancer Genome Interpreter (CGI)
(Tamborero et al., 2018). These harmonized variants are hosted on an
ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P,
\url{https://search.cancervariants.org/\#*}) that allows users to query
and filter aggregated evidence from the various databases listed above
as well as the GA4GH beacon service
(\url{http://beacon-network.org/\#/}), which yields access to genetic
mutation data from over 200 datasets as of 2016 (The Global Alliance for
Genomics and Health, 2016). For drug screen datasets, we focus on three
well-known, large-scale drug projects: the CCLE and GDSC, which were
mentioned above, and the Center for the Science of Therapeutics at the
Broad Institute's Cancer Therapeutics Response Portal (CTRP) dataset
(Basu et al., 2013; M. G. Rees et al., 2016; B. Seashore-Ludlow et al.,
2015). We also analyzed the results from a large-scale CRISPR/Cas9
screen conducted by the Broad Institute's Cancer Dependency Map project
(DepMap) using the Avana knockout library (Doench et al., 2016; Meyers
et al., 2017), which will address whether gene essentiality screens via
CRISPR yield results comparable to functional drug screens.

@@

After filtering, 894 level-A G2P interpretations were carried forward in
subsequent analyses, constituting 57 unique drugs and 34 unique genes in
88 distinct level-A combinations (\emph{Figure 1A: Pink heatmap}). The
number of level-A interpretations per drug ranged from 1 to 176 (\(M =\)
15.68, \(SD =\) 34.63), 1 to 264 (\(M =\) 26.24, \(SD =\) 61.58) per
gene, and for each level-A drug-gene combination, the range was 1 to 103
(\(M =\) 10.14, \(SD =\) 19.80). These unique drugs, genes, and
combinations are hereafter referred to as G2P drugs, G2P genes, and G2P
associations. Filtering our screening datasets by these clinical
criteria yields panels of potentially clinically relevant omic features.
Of the 57 unique G2P drugs, 37 of them were screened in at least one of
the drug screen datasets, 17 drugs were screened in at least two of the
three datasets, and only 3 were screened in all three datasets. The
remaining 20 un-screened G2P drugs suggests an incompleteness of these
drug screens, though these projects frequently release new screening
data that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P
associations tested in the CCLE, CTRP, and GDSC drug screens. Of the 88
G2P associations, data was available for 9 associations in CCLE, 48 in
CTRP, and 44 in GDSC. Controlling for a false discovery rate (FDR)
\textless{} 5\%, CCLE yielded no significant correlations (Wilcoxon
rank-sum tests, \(p_k = 1.2^{-4}\)), though the most significant G2P
association was selumetinib-\emph{KRAS} (\(p_{adj} = 1.1^{-3}\)). In
CTRP, AUC from 3 drug-gene associations correlated with the gene's
mutation status (\(p_k = 2.6^{-3}\)): vemurafenib-\emph{BRAF}
(\(p_{adj} = 5.8^{-9}\)), selumetinib-\emph{KRAS}
(\(p_{adj} = 4.6^{-8}\)), and dabrafenib-\emph{BRAF}
(\(p_{adj} = 8.3^{-5}\)). Dabrafenib-\emph{BRAF}
(\(p_{adj} = 3.2^{-11}\)) was the only significant association in GDSC
(\(p_k = 2.6^{-6}\)).

@@

To test for potentially novel drug-gene associations, we conducted
similar Wilcoxon rank-sum tests for all pairwise combinations of G2P
genes and drugs tested in the three drug screens (\emph{Figures 1B:
Green heatmaps}). At an unadjusted \(\alpha = 0.05\), the tests revealed
that in CCLE, AUC significantly correlated with mutation status for 14
out of 170 (8.24\%) tested drug-gene combinations were significantly
different, 93 out of 918 (10.13\%) were significantly different for
CTRP, and 72 out of 850 (8.47\%) for GDSC.

Controlling for a FDR \textless{} 5\%, selumetinib-\emph{BRAF}
(\(p_{adj} = 2.9^{-10}\)) was the only significant drug-gene combination
tested in CCLE (\(p_k = 1.2^{-4}\)). CTRP (\(p_k = 3.9^{-4}\)) yielded 4
significant combinations, one of which was selumetinib-\emph{BRAF}.
There were 2 significant G2P associations in the GDSC dataset
(\(p_k = 2.3^{-4}\)). Of all the significant correlations between
mutation status and AUC across datasets, the only genes represented were
\emph{KRAS} and \emph{BRAF} (Figure~\ref{fig:drug-venn}). Of the 5
significant drug-gene associations across datasets, 3 were G2P
associations. The remaining 2 associations were selumetinib-\emph{BRAF}
and trametinib-\emph{KRAS}.

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/drugscreen_venns.pdf}
\caption{\label{fig:drug-venn}\textbf{Overlap of significant drug-gene associations
from CCLE, CTRP, and GDSC}. Overlap of drug-gene associations drawn from
all pairwise combinations of G2P drugs and G2P genes whose drug effects
significantly correlated with (\textbf{A}) mutation status, (\textbf{B})
gene expression, and (\textbf{C}) copy number across CCLE, CTRP, and
GDSC.}
\end{figure}

@@

Prior to correction for an FDR of 5\%, 1 DGL,
selumetinib-\emph{KRAS}-lung cancer, was identified in the CCLE screen
when using mutation status as a correlate (\(n =\) 7, 2 DGLs were not
analyzed due to lack of mutation annotations in \emph{UGT1A1} in
\enquote{cancer} lineages), but CTRP (\(n =\) 48) and GDSC (\(n =\) 39)
identified 5 and 4, respectively, in skin cancer, lung cancer, and
cancer lineages. After Benjamini-Hochberg FDR correction, none of the
datasets yielded significant DGL interactions.

\begin{quote}
\textbf{(???) KS test results: Need to flesh out this section}
\end{quote}

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/g2p_spearman_ks_res_plot.pdf}
\caption{\label{fig:ksplot}\textbf{Comparison of drug-gene AUC vs.~gene expression and
vs.~copy number in G2P and non-G2P cell lines}. Comparison of Spearman
correlation coefficients (\(r_s\)) derived from copy number (log\(_2\)
ratio) vs.~AUC z-scores and gene expression (RPKM) vs.~AUC z-scores for
G2P drug-gene associations. Correlation coefficients were grouped by
whether or not the associated lineage for each comparison was in a
level-A G2P drug-gene-lineage association or not. The distributions of
these grouped coefficients were compared using Kolmogorov-Smirnov tests
to identify whether correlations in G2P lineages were distinct from
correlations in non-G2P lineages.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/crispr_spearman_corr_plot.pdf}
\caption{\label{fig:crispr-waterfall}Stuff stuff stuff!}
\end{figure}

\begin{figure}
\centering
\includegraphics{./plots/manuscript/crispr_venn.pdf}
\caption{\label{fig:crispr-venn}\textbf{CRISPR sceen results}. Overlap of genes whose
gene expression and/or copy number significantly correlated with CERES
score (Spearman rank-order correlations).}
\end{figure}

@@

When we don't account for lineage, Wilcoxon rank-sum tests
(\(p_k = 4.1^{-17}\), FDR \textless{} 5\%) revealed that mutations in 1
gene significantly correlated with CERES score: \emph{KRAS}
(\(p_{adj} = 3.7^{-37}\)). Gene expression (RPKM) of 13 out of 33
(39.39\%, Spearman rank-order correlations) G2P genes significantly
correlated with CERES score, as did copy number (log2-ratio) for 12 out
of 33 (36.36\%, Spearman rank-order correlations) G2P genes
(Figure~\ref{fig:crispr-waterfall}). We see moderate overlap between
these genes whose CERES scores significantly correlated with gene
expression and with copy number (Figure~\ref{fig:crispr-venn}). For the
purpose of comparison to the drug screen data, if we treat CRISPR
knockdown like a small-molecule drug, then these results are more
comprehensive than the results from CCLE, CTRP, and GDSC.

\begin{quote}
\textbf{(???) Lineage????}
\end{quote}

@@

In a lineage-agnostic context of the CRISPR screen, \emph{ERBB2} gene
essentiality score had significant negative correlations with
\emph{ERBB2} gene expression (\(p = 8.3^{-12}\), \(r_s =\) -0.30) and
\emph{ERBB2} copy number (\(p = 3.8^{-10}\), \(r_s =\) -0.27)
(\emph{Figure \#\#2}). Notably, gene essentiality of \emph{MET}, which
is known to have copy number alterations in some cancers (An et al.,
2014; J. H. Kim, Kim, \& Kim, 2018), did not significantly correlate
with \emph{MET} gene expression nor \emph{MET} copy number. \emph{BRCA1}
gene essentiality also significantly correlated with both \emph{BRCA1}
gene expression (\(p = 4.7^{-6}\), \(r_s =\) 0.20) and \emph{BRCA1} copy
number (\(p = 1.1^{-3}\), \(r_s =\) 0.14). \emph{BRCA2} gene
essentiality significantly correlated \emph{BRCA2} gene expression
(\(p = 6.6^{-7}\), \(r_s =\) 0.22) but not \emph{BRCA2} copy number.

@@

\begin{quote}
\textbf{(???) Should we do this?}
\end{quote}

@@

CRISPR/Cas9, CCLE mutation calls, and CCLE copy number data were all
retrieved from the Broad Institute's Cancer Dependency Map (DepMap)
Public 18Q4 release via the DepMap data portal
(\url{https://depmap.org/portal/download/}) (DepMap, 2018). Gene
expression data was retrieved from the DepMap Public 18Q3 release (Broad
DepMap, 2018) from the same data portal. For cell lines with no CCLE
annotation, the DepMap group drew raw copy number and mutation data from
whole exome sequencing data produced by the Wellcome Trust Sanger
Institute {[}Catalogue Of Somatic Mutations In Cancer (COSMIC,
\url{https://cancer.sanger.ac.uk/cell_lines}) (S. Bamford et al., 2004;
S. A. Forbes et al., 2017); European Genome-phenome Archive (Lappalainen
et al., 2015), accession number EGAD00001001039{]} and processed the
data following the CCLE pipelines to ensure consistency.

@@

Gene effect scores for 17,634 genes in 517 cell lines were inferred from
a CRISPR/Cas9 (clustered regularly interspaced short palindromic
repeats/CRISPR-associated 9) screen using the Broad Institute's Avana
knockout library (Doench et al., 2016). The Broad Institute's data
processing and screening methods are available from the figshare record
(DepMap, 2018) and the original publication of CERES, the algorithm that
computes inferred gene dependency scores (Meyers et al., 2017). The
DepMap releases new data quarterly, with the current data set being the
18Q4 release. This release has 175 more cell lines than the original
data release that was published with CERES (dataset: gene\_effect.csv,
accessed 15 November 2018).

@@

The CCLE generated genomic copy number (CN) data using the Affymetrix
Genome-Wide Human SNP Array 6.0 and GenePattern pipeline (Network, 2008)
and normalized segmented CN log2-ratios for 23,299 genes across 1,098
cell lines using circular binary segmentation (Olshen, Venkatraman,
Lucito, \& Wigler, 2004) (dataset: public\_18Q4\_gene\_cn.csv, accessed
15 November 2018).

@@

For \emph{Figure 1}, all data were imported and manipulated using pandas
(v0.23.0) (McKinney, 2011) and plotted using seaborn (v0.90) (Waskom,
2012), which sat on top of matplotlib (v2.2.2) (Hunter, 2007) and was
executed on Python (v2.7.15) (Python Software Foundation,
\url{https://www.python.org/}) in an IPython (v5.5.0) (Pérez \& Granger,
2007) kernel within a localhost Jupyter (v5.2.2) notebook (Kluyver et
al., 2016). All other analyses were conducted in R (v3.5.0) (R. D. C.
Team, 2010, \url{https://www.R-project.org/}) in an RStudio environment
(v1.1.447) (RStudio, 2012). All other plots were drawn using the ggplot2
(v3.0.0) R package (Wickham, 2016).

@@

For computing Spearman correlation coefficients, we used the
\texttt{cor.test} function from the \texttt{stats} package. If there
were 3 or fewer data points, then the correlation was not done and a
value of \texttt{NA} was supplied.

@@

\hypertarget{ref-an2014}{}
An, X., Wang, F., Shao, Q., Wang, F.-H., Wang, Z.-Q., Chen, C., \ldots{}
Li, Y.-H. (2014). MET amplification is not rare and predicts unfavorable
clinical outcomes in patients with recurrent/metastatic gastric cancer
after chemotherapy. \emph{Cancer}, \emph{120}(5), 675--682.
doi:\href{https://doi.org/10.1002/cncr.28454}{10.1002/cncr.28454}

@@

@@

\hypertarget{ref-broaddepmap2018a}{}
DepMap, B. (2018, January 11). DepMap Achilles 18Q4 public. Figshare.
Retrieved from
\url{https://figshare.com/articles/DepMap_Achilles_18Q4_public/7270880}

@@

\hypertarget{ref-kim2018}{}
Kim, J. H., Kim, H. S., \& Kim, B. J. (2018). Prognostic value of MET
copy number gain in non-small-cell lung cancer: An updated
meta-analysis. \emph{J Cancer}, \emph{9}(10), 1836--1845.
doi:\href{https://doi.org/10.7150/jca.24980}{10.7150/jca.24980}

@@

\hypertarget{ref-thecancergenomeatlasresearchnetwork2008}{}
Network, T. C. G. A. R. (2008). Comprehensive genomic characterization
defines human glioblastoma genes and core pathways, \emph{455},
1061--1068.
doi:\href{https://doi.org/10.1038/nature07385}{10.1038/nature07385}

@@

@@

\hypertarget{ref-rdevelopmentcoreteam2010}{}
Team, R. D. C. (2010). \emph{R: A Language and Environment for
Statistical Computing}. Vienna, Austria: R Foundation for Statistical
Computing. Retrieved from \url{https://www.R-project.org/}

### 2019-01-24 manuscript_analysis.Rmd
updated manuscript_analysis and manuscript

crispr_ge_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = RPKM, y = Score)) +
  geom_point(pch = 21, fill = "cadetblue3", alpha = 0.6) +
  labs(x = "KRAS Gene Expression (RPKM)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[1], 2), " (", printp(spearman_KRAS$p[1]), ")"))

crispr_cn_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = CN, y = Score)) +
  geom_point(pch = 21, fill = "palegreen2", alpha = 0.6) +
  labs(x = "KRAS Copy Number (log2 ratio)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[2], 2), " (", printp(spearman_KRAS$p[2]), ")"))

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "deepskyblue4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_ge_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 3.5, y = -0.27, xend = 14, yend = -0.165), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

@@

crispr_spearman_corr_plot <- plot_grid(crispr_ge_spearman_corr_plot, crispr_cn_spearman_corr_plot, align = "h")
crispr_spearman_corr_plot
# ggsave("./plots/manuscript/crispr_spearman_corr_plot.png", crispr_spearman_corr_plot, device = "png", dpi = 450, width = 16, height = 8, units = "in")
```

>>>>

crispr_ge_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = RPKM, y = Score)) +
  geom_point(pch = 21, fill = "cadetblue3", alpha = 0.6) +
  labs(x = "KRAS Gene Expression (RPKM)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[1], 2), " (p = ", formatC(as.numeric(spearman_KRAS$p[1]), format = "e", digits = 1), ")"))

crispr_cn_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = CN, y = Score)) +
  geom_point(pch = 21, fill = "palegreen2", alpha = 0.6) +
  labs(x = "KRAS Copy Number (log2 ratio)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[2], 2), " (p = ", formatC(as.numeric(spearman_KRAS$p[2]), format = "e", digits = 1), ")"))

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "deepskyblue4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = expression("Spearman Correlation Coefficient"~(r[s]))) +
  annotation_custom(ggplotGrob(crispr_ge_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 3.5, y = -0.27, xend = 14, yend = -0.165), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

@@

crispr_spearman_corr_plot <- plot_grid(crispr_ge_spearman_corr_plot, crispr_cn_spearman_corr_plot, align = "h")
crispr_spearman_corr_plot

# ggsave("./plots/manuscript/crispr_spearman_corr_plot.pdf", crispr_spearman_corr_plot, device = "pdf", dpi = 450, width = 16, height = 8, units = "in")
```
